---
title: "Geographic Mapping with tuikr"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Geographic Mapping with tuikr}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE,
  fig.width = 8,
  fig.height = 6
)
```

## Introduction

The `tuikr` package provides access to geographic statistical data from TUIK's Geographic Statistics Portal. This vignette demonstrates how to:

1. Discover available geographic variables
2. Download data at different NUTS levels
3. Create choropleth maps
4. Join statistical data with spatial boundaries
5. Create advanced cartographic visualizations

## Understanding NUTS Levels in Turkey

Turkey's administrative divisions follow the NUTS (Nomenclature of Territorial Units for Statistics) system:

| Level | Turkish Term | English Translation | Count | Example |
|-------|--------------|---------------------|-------|---------|
| **NUTS-2** | İstatistiki Bölge | Statistical Region | 26 | İstanbul, Ankara |
| **NUTS-3** | İl | Province | 81 | İzmir, Antalya |
| **LAU-1** | İlçe | District | 973 | Kadıköy, Çankaya |
| **Level 9** | Yerleşim Noktası | Settlement Point | 1,003 | City centers |

**Key Differences:**
- Levels 2, 3, and 4 return `MULTIPOLYGON` geometry (areas)
- Level 9 returns `POINT` geometry (specific locations)

## Setup

```{r setup}
library(tuikr)
library(tidyverse)
library(sf)
```

## Discovering Available Variables

Start by exploring what geographic data is available:

```{r variables}
# Get all available variables
variables <- geo_data()

# View the structure
glimpse(variables)

# Example: Search for education-related data
variables %>%
  filter(str_detect(var_name, "Okul|Eğitim"))
```

The `geo_data()` function without parameters returns metadata for all available variables:

- `var_name`: Variable name (Turkish)
- `var_num`: Variable code (unique identifier)
- `var_levels`: Available NUTS levels (list column)
- `var_period`: Data period type (yillik = annual)
- `var_source`: Data source name
- `var_recordnum`: Number of records

## Downloading Geographic Data

### Basic Download

To download data for a specific variable:

```{r download}
# Example: Cinema film count at NUTS-3 level
cinema_data <- geo_data(
  variable_level = 3,
  variable_no = "SNM-GK160951-O33303",
  variable_source = "medas",
  variable_period = "yillik",
  variable_recnum = 5
)

head(cinema_data)
```

### Finding Variable Parameters

You can extract the necessary parameters from the metadata:

```{r params}
# Find a variable of interest
var_info <- variables %>%
  filter(var_num == "SNM-GK160951-O33303")

# Extract parameters
var_info %>%
  select(var_num, var_source, var_period, var_recordnum, var_levels)

# Check available levels
var_info$var_levels[[1]]  # Returns: c(3, 4)
```

### Level Availability

Not all variables are available at all NUTS levels. If you request an unavailable level:

```{r level_error, error=TRUE}
geo_data(
  variable_level = 4,
  variable_no = "TFE-GK105747-O23001",
  variable_source = "medas",
  variable_period = "yillik",
  variable_recnum = 5
)
#> Error: This data (TFE-GK105747-O23001) is not available at this NUTS level (level = 4)!!!
```

## Downloading Map Boundaries

Download spatial boundaries for mapping:

```{r boundaries}
# NUTS-2 (26 regions)
nuts2 <- geo_map(2)

# NUTS-3 (81 provinces)
nuts3 <- geo_map(3)

# LAU-1 (973 districts)
lau1 <- geo_map(4)

# Settlement points
settlements <- geo_map(9)
```

Map objects are `sf` (simple features) objects with:
- **CRS**: WGS 84 (EPSG:4326)
- **Columns**:
  - `code`: Numeric identifier for joining
  - `bolgeKodu`: NUTS-2 region code
  - `nutsKodu`: NUTS code
  - `name`, `ad`: Place name (English/Turkish)
  - `geometry`: Spatial geometry

## Creating Choropleth Maps

### NUTS-2 Example: Chicken Population

```{r nuts2_map}
# 1. Find the variable
variables %>%
  filter(var_num == "HYV-GK1696800-O32507")

# 2. Download data for 2023 and 2015
chicken <- geo_data(
  variable_level = 2,
  variable_no = "HYV-GK1696800-O32507",
  variable_source = "medas",
  variable_period = "yillik",
  variable_recnum = 20
) %>%
  filter(date %in% c("2023", "2015")) %>%
  mutate(
    yumurta_tavugu_sayisi_adet = as.numeric(yumurta_tavugu_sayisi_adet),
    date = as.numeric(date)
  )

# 3. Join with spatial data and map
geo_map(2) %>%
  left_join(chicken, by = "code") %>%
  ggplot() +
  geom_sf(aes(fill = yumurta_tavugu_sayisi_adet), color = "white") +
  coord_sf(datum = NA) +
  scale_fill_viridis_c(option = "plasma") +
  theme_minimal() +
  theme(
    legend.position = "bottom",
    legend.key.width = unit(3, "cm")
  ) +
  labs(
    fill = "Adet",
    title = "Yumurta Tavuğu Sayısı",
    caption = "Kaynak: TÜİK"
  ) +
  facet_wrap(~date, ncol = 2)
```

### NUTS-3 Example: Housing Sales

```{r nuts3_map}
# 1. Find housing sales data
variables %>%
  filter(var_num == "INS-GK055-O006")

# 2. Download 2019 data
house <- geo_data(
  variable_level = 3,
  variable_no = "INS-GK055-O006",
  variable_source = "ilGostergeleri",
  variable_period = "yillik",
  variable_recnum = 5
) %>%
  filter(date == 2019) %>%
  mutate(konut_satis_sayilari_toplam = as.numeric(konut_satis_sayilari_toplam))

# 3. Create map
geo_map(3) %>%
  left_join(house, by = "code") %>%
  ggplot() +
  geom_sf(aes(fill = konut_satis_sayilari_toplam)) +
  coord_sf(datum = NA) +
  scale_fill_viridis_c(option = "magma") +
  theme_minimal() +
  theme(
    legend.position = "bottom",
    legend.key.width = unit(3, "cm"),
    plot.title = element_text(hjust = 0.5)
  ) +
  labs(
    fill = "Satış Sayısı",
    title = "2019 Yılında Konut Satış Sayıları",
    caption = "Kaynak: TÜİK"
  )
```

### LAU-1 Example: Illiteracy Rates

```{r lau1_map}
# 1. Download illiteracy data at district level
illiteracy <- geo_data(
  variable_level = 4,
  variable_no = "ULE-GK160887-O29502",
  variable_source = "medas",
  variable_period = "yillik",
  variable_recnum = 5
) %>%
  filter(date == 2022) %>%
  mutate(okuma_yazma_bilmeyen_sayisi = as.numeric(okuma_yazma_bilmeyen_sayisi))

# 2. Join with district boundaries
geo_map(4) %>%
  left_join(illiteracy, by = "code") %>%
  ggplot() +
  geom_sf(aes(fill = okuma_yazma_bilmeyen_sayisi), lwd = 0.1) +
  coord_sf(datum = NA) +
  scale_fill_viridis_c(option = "inferno") +
  theme_minimal() +
  theme(
    legend.position = "bottom",
    legend.key.width = unit(3, "cm"),
    plot.title = element_text(hjust = 0.5)
  ) +
  labs(
    fill = "Sayı",
    title = "2022 Yılında Okuma Yazma Bilmeyen Sayısı",
    caption = "Kaynak: TÜİK"
  )
```

### Regional Focus: Ankara Region

```{r regional}
# Focus on Ankara region (TR51)
population <- geo_data(
  variable_level = 4,
  variable_no = "ADNKS-GK137473-O29001",
  variable_source = "medas",
  variable_period = "yillik",
  variable_recnum = 5
) %>%
  filter(date == 2019) %>%
  mutate(toplam_nufus = as.numeric(toplam_nufus))

geo_map(4) %>%
  left_join(population, by = "code") %>%
  filter(bolgeKodu == "TR51") %>%
  ggplot() +
  geom_sf(aes(fill = toplam_nufus)) +
  scale_fill_viridis_c(option = "viridis") +
  theme_minimal() +
  labs(
    fill = "Nüfus",
    title = "Ankara Bölgesi İlçe Nüfusları (2019)",
    caption = "Kaynak: TÜİK"
  )
```

## Advanced Cartography

### Hexagonal Cartograms

Hexagonal maps provide equal visual weight to each province:

```{r hex, message=FALSE}
library(geogrid)

# 1. Transform CRS for better projection
tur_hex_map <- st_transform(geo_map(3), crs = 3395)

# 2. Download and join automobile data
auto_data <- geo_data(
  variable_level = 3,
  variable_no = "ULS-GK093-O009",
  variable_source = "ilGostergeleri",
  variable_period = "yillik",
  variable_recnum = 5
) %>%
  filter(date == 2019) %>%
  mutate(bin_kisi_basina_otomobil_sayisi = as.numeric(bin_kisi_basina_otomobil_sayisi))

tur_hex_dt <- left_join(tur_hex_map, auto_data, by = "code")

# 3. Calculate hexagonal grid
new_hex <- calculate_grid(
  shape = tur_hex_dt,
  grid_type = "hexagonal",
  seed = 9
)
result_hex <- assign_polygons(tur_hex_dt, new_hex)

# 4. Plot
result_hex %>%
  mutate(
    name = str_replace(name, "AFYONKARAHİSAR", "AFYON"),
    name = str_replace(name, "KAHRAMANMARAŞ", "K.MARAŞ")
  ) %>%
  ggplot() +
  geom_sf(aes(fill = bin_kisi_basina_otomobil_sayisi), lwd = 0.1) +
  geom_sf_text(aes(label = name), color = "black", size = 2) +
  coord_sf(datum = NA) +
  scale_fill_viridis_c(option = "plasma") +
  theme_minimal() +
  theme(
    legend.position = "bottom",
    legend.key.width = unit(3, "cm")
  ) +
  labs(
    fill = "Oran",
    title = "Bin Kişi Başına Otomobil Sayısı (2019)",
    caption = "Kaynak: TÜİK"
  )
```

### Dorling Cartograms

Dorling cartograms use circles sized by a variable, preserving geographic relationships while emphasizing data values:

```{r dorling, message=FALSE}
library(cartogram)

# 1. Download population and illiteracy data
tur_map <- st_transform(geo_map(4), crs = 3395)

population <- geo_data(
  variable_level = 4,
  variable_no = "ADNKS-GK137473-O29001",
  variable_source = "medas",
  variable_period = "yillik",
  variable_recnum = 5
) %>%
  filter(date == 2019) %>%
  mutate(toplam_nufus = as.numeric(toplam_nufus))

illiteracy <- geo_data(
  variable_level = 4,
  variable_no = "ULE-GK160887-O29502",
  variable_source = "medas",
  variable_period = "yillik",
  variable_recnum = 5
) %>%
  filter(date == 2019) %>%
  mutate(okuma_yazma_bilmeyen_sayisi = as.numeric(okuma_yazma_bilmeyen_sayisi))

# 2. Combine datasets
combined <- left_join(population, illiteracy, by = c("code", "date")) %>%
  left_join(tur_map, ., by = "code") %>%
  mutate(illiteracy_rate = 100 * (okuma_yazma_bilmeyen_sayisi / toplam_nufus))

# 3. Create Dorling cartogram (sized by illiteracy count)
dorling <- cartogram_dorling(combined, "okuma_yazma_bilmeyen_sayisi", 0.4)

# 4. Plot
dorling %>%
  ggplot() +
  geom_sf(aes(fill = illiteracy_rate), lwd = 0.1) +
  coord_sf(datum = NA) +
  scale_fill_viridis_c(option = "inferno") +
  theme_minimal() +
  theme(
    legend.position = "bottom",
    legend.key.width = unit(3, "cm")
  ) +
  labs(
    fill = "Oran (%)",
    title = "Okuma Yazma Bilmeyen Oranı (2019)",
    subtitle = "Daire boyutu: toplam sayı",
    caption = "Kaynak: TÜİK"
  )
```

## Tips for Working with Geographic Data

### 1. Always Convert to Numeric

Geographic data downloads return character columns. Convert to numeric for analysis:

```{r convert}
data <- geo_data(3, "SNM-GK160951-O33303", "medas", "yillik", 5) %>%
  mutate(across(where(is.character) & !c(code, date), as.numeric))
```

### 2. Join by Code

Use the `code` column for joining spatial and statistical data:

```{r join}
geo_map(3) %>%
  left_join(statistical_data, by = "code")
```

### 3. Handle Date Formats

Date columns may be in `YYYY` or `YYYYMM` format:

```{r dates}
# Annual data: "2019"
# Monthly data: "201901" (January 2019)

data %>%
  mutate(
    year = as.numeric(substr(date, 1, 4)),
    month = if_else(
      nchar(date) == 6,
      as.numeric(substr(date, 5, 6)),
      NA_real_
    )
  )
```

### 4. Remove Coordinate Grid Lines

For cleaner maps, use `coord_sf(datum = NA)`:

```{r clean}
ggplot(map_data) +
  geom_sf(aes(fill = value)) +
  coord_sf(datum = NA)  # Removes lat/long grid
```

### 5. Transform CRS for Cartograms

For cartogram functions, transform to a projected CRS:

```{r transform}
# WGS 84 / World Mercator
map_projected <- st_transform(map_data, crs = 3395)
```

## Next Steps

- Explore all available variables: `geo_data()`
- Check the package documentation: `?geo_data`, `?geo_map`
- For known issues and limitations, see `vignette("known-issues")`
- Share your maps on social media with #tuikr!

## Further Reading

- **sf package**: [r-spatial.github.io/sf](https://r-spatial.github.io/sf/)
- **geogrid package**: [CRAN](https://cran.r-project.org/package=geogrid)
- **cartogram package**: [CRAN](https://cran.r-project.org/package=cartogram)
- **TUIK Geographic Portal**: [cip.tuik.gov.tr](https://cip.tuik.gov.tr/)
